DATA 621 Homework #4

### 2019-11-14 hw4/HW4.Rmd
Minor change
Update HW4.Rmd
Fixed misc model and added OLS models

There are `r sum(train$TARGET_FLAG)` out of `r nrow(train)` records in the training data set that have been in an accident.  We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.

@@

## Model Creation

@@

```{r, null_prefix = TRUE, comment=NA}
#"KIDSDRIV"         "AGE"             
#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
#[36] "KID_DRIVERS"     
model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
             family = binomial(link = "logit"), over_sample_train)
summary(model)

@@

>>>>

There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.  We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.

@@

## Model Creation - Logistic

@@

Here we add meaningful variables to the model until we have a robust model with a decent AUC value.

```{r, null_prefix = TRUE, comment=NA}
model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + LOG_INCOME +
             SEX + CAR_USE + CAR_AGE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +
               TIF + TRAVTIME + AGE,
             family = binomial(link = "logit"), over_sample_train)
summary(model)

@@

```{r}
# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(model,newdata=test,type="response"), prob = 
predict(model,newdata=test,type="response"))

pred <- prediction(eval$predicted, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

## Model Creation - Linear

First we need to remove the zero values from our train and test sets, as we are only interested in cases where there was an accident claim.

```{r}
train <- train %>% filter(TARGET_AMT > 0)
test <- test %>% filter(TARGET_AMT > 0)
```

As with the logistic model, we start with a baseline model using the `BLUEBOOK` variable as the sole predictor. All other models will need to beat this baseline.

```{r, null_prefix = TRUE, comment=NA}

baseline_model <- lm(TARGET_AMT ~ BLUEBOOK, train)
summary(baseline_model)
```

This model is rather poor, though the predictor is statistically significant in the model. We can build on this model and see if we can build a more robust version.

### Car Model

Here we look at the type of car the person drives, it's age, and how it is used.

```{r}
car_model <- lm(TARGET_AMT ~ CAR_AGE + CAR_TYPE + CAR_USE, data = train)
summary(car_model)
```

Here we see that none of these predictors are significant in this model. In fact, taken one at a time, none are significant individually either.

### Personal Model

Here we use variables pertaining to the driver's personal life: family, income, driving record, and work.

```{r}
personal_model <- lm(TARGET_AMT ~ MSTATUS + INCOME + KIDSDRIV +
                       HOME_VAL + AGE + REVOKED + MVR_PTS +
                       HOMEKIDS, data= train)

summary(personal_model)
```

Again, we see a very poor model. No variables are statistically significant here (though 2 are close).

### Additive Model

Here we start with `BLUEBOOK` from the baseline model and add to it until we get a decent model.

```{r}
add_model <- lm(TARGET_AMT ~ BLUEBOOK + CLM_FREQ, data = train)

summary(add_model)
```
